---
title: "Age-Labor Income Profile"
author: "Sany León, Andrés Suárez, and Juan Rueda"
date: "`r Sys.Date()`"
output:
  beamer_presentation: default
  slidy_presentation: default
institute: Universidad de Los Andes
subtitle: Predicting Income
theme: default
fonttheme: serif
header-includes: 
  - \useoutertheme{infolines}
  - \usepackage{booktabs}
  - \usepackage{graphicx}

---

#############################################################################
# **Research Question**

## ¿How does wage vary with age in Bogota?

\vspace{0.5cm}
In Bogota, wages tend to increase with age until a mid-life point and then decrease.  

\vspace{0.5cm}
This is consistent with the **human capital theory**:  

- Wages increase at a decreasing rate.

#############################################################################

# **Data**
Estadísticas descriptivas relevantes

#############################################################################

# Results

A regression table that demonstrates the contribution of the quadratic term to model fit and the effect of including controls on the estimated age–income relationship. The table should report the implied peak age (with confidence intervals) and in-sample fit for each specification presented.

\begin{center}
\includegraphics[width=0.5\textwidth]{"../02_outputs/figures/age_income_profiles.png"}
\end{center}


#############################################################################

# Results

```{r results='asis', echo=FALSE}
cat("\\begin{center}\n")
cat("\\tiny\n")   # o \\scriptsize o \\tiny
cat(readLines("../02_outputs/tables/02_model_age_income_peak.tex"), sep="\n")
cat("\n\\end{center}\n")
```

#############################################################################

# Results

```{r results='asis', echo=FALSE}
cat("\\begin{center}\n\\scriptsize\n")
cat(readLines("../02_outputs/tables/03_model_gender_gap.tex"), sep="\n")
cat("\n\\end{center}\n")
```


#############################################################################

# **Discussion**
La tabla muestra la relación edad-salario con modelos lineales y cuadráticos, incluyendo controles laborales y tipo de empleo.  
El modelo lineal simple no captura el patrón no lineal (\(R^2 = 0.004\)).  
Al incluir el término cuadrático, el salario alcanza un máximo alrededor de 40 años.  
Con controles laborales, el pico ajusta a 41 años y mejora el ajuste (\(R^2 = 0.239\)).  
Los coeficientes de tipo de empleo indican diferencias significativas respecto a la categoría base.
