---
title: "**Age-Labor Income Profile**"
author: "Sany León, Andrés Suárez, and Juan Rueda"
date: "`r Sys.Date()`"
output:
  beamer_presentation: 
  slidy_presentation: default
#institute: 
subtitle: Universidad de Los Andes
theme: default
fonttheme: serif
fontsize: 10pt
header-includes: 
  - \useoutertheme{infolines}
  - \usepackage{booktabs}
  - \usepackage{graphicx}
  - \usepackage{threeparttable}
  - \usepackage{booktabs}
  - \usepackage{longtable}
---

# **Research Question**


To what extent do labor income models capture the structural determinants of income, and how does the balance between their interpretability and predictive accuracy affect their usefulness for policymaking?

\vspace{0.5cm}
Is there empirical evidence of a non-linear (concave) relationship between age and labor income in Bogotá, and how does this relationship change when controlling for hours worked and type of employment?

#############################################################################

# **Data**
```{r results='asis', echo=FALSE}
cat("\\scriptsize\n")
cat(readLines("../02_outputs/tables/01_tabla_continuas.tex"), sep="\n")
```



#############################################################################

## **Data**
```{r results='asis', echo=FALSE}
cat("\\scriptsize\n")
cat(readLines("../02_outputs/tables/01_tabla_categoricas.tex"), sep="\n")
```


#############################################################################

# **Results**

\begin{figure}[h!]
\centering
\caption{Descriptive Statistic Continuos variables}
\includegraphics[width=0.6 \textwidth]{"../02_outputs/figures/age_income_profiles.png"}
\label{fig:age_income}
\end{figure}

#############################################################################

# **Results**


```{r results='asis', echo=FALSE}
cat("\\begin{center}\n")
cat("\\tiny\n")   # o \\scriptsize o \\tiny
cat(readLines("../02_outputs/tables/02_model_age_income_peak.tex"), sep="\n")
cat("\n\\end{center}\n")
```


#############################################################################

# **Discussion**

\vspace{0.5cm}
### 1. Evidence of a Non-Linear and Concave Age–Income Profile

\vspace{0.5cm}
The quadratic specifications show a positive coefficient on age and a negative coefficient on age², generating a **concave age–earnings profile**. There is clear evidence of an **earnings peak within the observed age range**, around 40 years.

\vspace{0.5cm}
In contrast, the linear model fails to capture this non-linearity and exhibits very low explanatory power.

\vspace{0.5cm}
- Results are consistent with the theory: "Earnings rise at a diminishing rate over the working life, and decline when net investment becomes negative, as in old age. The typical (logarithmic) working-life earnings profile is, therefore, concave" (Mincer, 1974)

#############################################################################

# **Discussion**

### 2. Effect of Conditioning on Hours Worked and Employment Type

Including controls for hours worked and employment type shifts the estimated **earnings peak to slightly older ages (approximately 41–43 years)** and substantially improves model fit.  

This suggests that part of the observed non-linearity is driven by differences in labor intensity and occupational structure, rather than solely by human capital accumulation.

### 3. Economic Interpretation in the Context of Bogotá’s Labor Market

The results support the theoretical prediction of a concave life-cycle earnings profile. However, the shape of the profile is influenced by labor market characteristics such as occupational heterogeneity and variation in working hours.  

Overall, the observed age–income pattern reflects both experience accumulation and structural labor market factors.

#############################################################################

# **References**

Mincer, J. A. (1974). Schooling, Experience, and Earnings. National Bureau of Economic Research.
