---
title: "**Age-Labor Income Profile**"
author: "Sany León, Andrés Suárez, and Juan Rueda"
date: "`r Sys.Date()`"
output:
  beamer_presentation: 
  slidy_presentation: default
#institute: 
subtitle: Universidad de Los Andes
theme: default
fonttheme: serif
fontsize: 10pt
header-includes: 
  - \useoutertheme{infolines}
  - \usepackage{booktabs}
  - \usepackage{graphicx}
  - \usepackage{threeparttable}
  - \usepackage{booktabs}
 # - \usepackage{longtable}
---

# **Research Question**


To what extent do labor income models capture the structural determinants of income, and how does the balance between their interpretability and predictive accuracy affect their usefulness for policymaking?

\vspace{0.5cm}
Is there empirical evidence of a non-linear (concave) relationship between age and labor income in Bogotá, and how does this relationship change when controlling for hours worked and type of employment?

#############################################################################

# **Data**
```{r results='asis', echo=FALSE}
cat("\\begin{table}[H]\n")
cat("\\centering\n")
cat("\\caption{Summary statistics for continuous variables}\n")
cat("\\label{tab:desc_continuous}\n")
cat("\\scriptsize\n")
cat(readLines("../02_outputs/tables/01_tabla_continuas.tex"), sep="\n")
cat("\n\\end{table}\n")
```



#############################################################################

## **Data**
```{r results='asis', echo=FALSE}
cat("\\begin{table}[H]\n")
cat("\\centering\n")
cat("\\caption{Distribution of categorical variables}\n")
cat("\\label{tab:desc_categorical}\n")
cat("\\scriptsize\n")
cat(readLines("../02_outputs/tables/01_tabla_categoricas.tex"), sep="\n")
cat("\n\\end{table}\n")
```


#############################################################################

# **Results**

\begin{figure}[h!]
\centering
\caption{Descriptive Statistic Continuos variables}
\includegraphics[width=0.6 \textwidth]{"../02_outputs/figures/age_income_profiles.png"}
\label{fig:age_income}
\end{figure}

#############################################################################

# **Results**


```{r results='asis', echo=FALSE}
cat("\\begin{center}\n")
cat("\\tiny\n")   # o \\scriptsize o \\tiny
cat(readLines("../02_outputs/tables/02_model_age_income_peak.tex"), sep="\n")
cat("\n\\end{center}\n")
```


#############################################################################

# **Discussion**

\vspace{0.5cm}
### 1. Evidence of a Non-Linear and Concave Age–Income Profile

\vspace{0.5cm}
The quadratic specifications show a positive coefficient on age and a negative coefficient on age², generating a **concave age–earnings profile**. There is clear evidence of an **earnings peak within the observed age range**, around 40 years.

\vspace{0.5cm}
In contrast, the linear model fails to capture this non-linearity and exhibits very low explanatory power.

\vspace{0.5cm}
- Results are consistent with the theory: "Earnings rise at a diminishing rate over the working life, and decline when net investment becomes negative, as in old age. The typical (logarithmic) working-life earnings profile is, therefore, concave" (Mincer, 1974)

#############################################################################

# **Discussion**

### 2. Effect of Conditioning on Hours Worked and Employment Type

\vspace{0.5cm}
Including controls for hours worked and employment type shifts the estimated **earnings peak to slightly older ages (approximately 41–43 years)** and substantially improves model fit.  

\vspace{0.5cm}
This suggests that part of the observed non-linearity is driven by differences in labor intensity and occupational structure, rather than solely by human capital accumulation.

\vspace{0.5cm}
- Including hours worked as a control is crucial, since “any analysis of labor supply requires an understanding of … hours of work” (Blundell & MaCurdy, 1999).


#############################################################################

# **Discussion**

### 3. Economic Interpretation in the Context of Bogota’s Labor Market

\vspace{0.5cm}
The results support the theoretical prediction of a concave life-cycle earnings profile. However, the shape of the profile is influenced by labor market characteristics such as occupational heterogeneity and variation in working hours.  

\vspace{0.5cm}
Overall, the observed age–income pattern reflects both experience accumulation and structural labor market factors.

\vspace{0.5cm}
- This result is consistent with Bogota’s labor-market structure, where informality is substantial: “Medellín or Bogotá, have [informality] rates of about 50%” (García, 2013).

#############################################################################

# **References**

Mincer, J. A. (1974). Schooling, Experience, and Earnings. National Bureau of Economic Research.

Blundell, R., & MaCurdy, T. (1999). Labor supply: A review of alternative approaches. In O. Ashenfelter & D. Card (Eds.), Handbook of Labor Economics (Vol. 3A). Elsevier.

García, G. A. (2013). Labor informality: choice or sign of segmentation? A Quantile Regression Approach at the Regional Level for Colombia (Preliminary version).
